---
title: "Lab14"
author: "Lil Rascals"
date: "April 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
library(gapminder)
library(modelr)
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
```

### Questions/Importance:
# 1) How does smoking affect the number of premature babies that are born?
This question is very important because it can point to whether or not smoking affects newborns, and if it does, how drastically to affects them. If our findings can give strong evidence toward smoking causing an increased chance of a premature baby, then we could argue that smoking puts the life of your baby at a much higher risk of death. Then we can also use it to advise against smoking during pregnancy.

# 2) What is the relationship between babies weight and whether the mother smokes accounting for a mother's weight during the pregnancy?
This is an interesting question because the answer could help us to find a better understand of how smoking during pregnancy impacts one's child. This is important because knowing this may help to discourage people from continuing to smoke or starting smoking, and encourage those who want to improve their babies well-being. Especially, in regards to mothers and how it impacts their children.



### Team Plots:
```{r}
babies1 <- babies %>% filter(Premature=="1")
ggplot(data = babies1) + 
  geom_jitter(mapping = aes(x = mpregwt, y = bwtoz))+
  geom_smooth(mapping = aes(x = mpregwt, y = bwtoz))+
  ggtitle("Mothers Weight vs. Child's Weight for Premature")
ggplot(data = babies) +
    geom_bar(mapping=aes(x=number, fill = as.factor(Premature)),position = "dodge") + 
    ggtitle("Number of Births vs. Number of Cigarettes Smoked Per Day") +
    xlab("Cigarettes Smoked per Day") +
    ylab("Number of Births")+
    scale_fill_discrete(name = "             Premature births 
(0 = not premature, 1 = premature)")
```

```{r}
ggplot(data = babies) + 
  geom_jitter(mapping = aes(x = mpregwt, y = bwtoz, color = smoke))+
  geom_smooth(mapping = aes(x = mpregwt, y = bwtoz, color = smoke))
```

### Our Findings:
1) Our findings were that generally the population of women who smoked the most had the fewest babies, and that the proportion of premature to mature born babies was roughly the same for women who didn't smoke and women who smoked multiple cigarettes a day historically. 
By adding a plot after filtering only the premature babies, we were able to see the birthweight vs. the mother's birthweight for prematurely born babies and see that it is generally upward sloping. 

2) Our team saw that from our plot that women who had a heigher pregnancy weight, had babies with a higher weight which makes sense. Also, we can tell from this plot that babies with mothers who smoked were less likely to have a higher weight and more likely to be a lower weight. Therefore, this could indicate that babies of mothers who smoke are generally less healthy. 

### Our Conclusions:
1) The data shows that women who smoked more cigarettes did have a larger proportion of premature babies, although women who smoked more in general had fewer babies overall.  
Also, we could see that of the premature babies, the birthweight was likely to be higher if the mother's birthweight was higher. 
2) The data shows that smoking of a mother, leads to a less normal baby weight. This could mean that these babies are less healthy or are more likely to be underweight. 

### Our Reccomendations: 
1) Our recommendation is that further study and analysis is required to produce a more confident conclusion on whether smoking during a pregnancy affects if a baby is born prematurely. However, smoking is definitively injurious to health and this data does show an increased proportion of premature babies born to women who have historically smoked several cigarettes. In general, women should be advised that smoking can affect the birth of a baby and should plan accordingly while further study continues. 

2) Our reccomendation is that we should continue to analyze more data specifcally trying to learn more about the problem of mothers smoking during or before pregnancy. There should be more legislation put in place to help us to combat the negative impact of smoking especially in regards to mothers. 

### Individual Plots:

#### Kelsey's PLot

```{r}

babies3 <- babies %>% filter(smoke > 0)
a <-mean(babies3$bwtoz)
b <- mean(babies$bwtoz)
c <- sd(babies3$bwtoz)
zscore <- (a-b)/c
zscore
ggplot(data = babies3) + 
  geom_point(mapping = aes(x = number, y = bwtoz, color = smoke))+
  xlab("Number of Cigarettes Smoked") + 
  ylab("Weight of Baby")

```
Kelsey's Question: How does whether a mother smokes affect her child's birthweight?

Kelsey's Findings: I found that there is very little relationship between how many cigarettes smoked a day and a babies birthweight. I found that there are more higher weighted babies as outliers if the mother doesn't smoke.
I took the weight of babies with mothers who smoked and then the rest of the population and compared them. I found that the probability of the variation by chance was like 80%, therefore, we cannot reject the probability that the means differ by chance alone. So our results are not statistically significant. 
I was able to apply filtering and testing through a difference of means test!


#### James' Plot 

James' Question: Does income affect the proportion of those who smoke, and those who do not smoke?


```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
ggplot(data = babies) +
  geom_bar(mapping = aes(x = inc, fill = as.factor(smoke)), position = position_dodge()) +
  xlab("Income")+
  scale_fill_discrete(name = "    Smoking Status
(0 = never, 1 = smokes,
NA = smoked before)")

```


James' Findings: I found that for the most part a families income does not affect whether or not they smoke. For most income groups the proportion of people who never smoke is greater than that of the proportion of people who smoke.


#### Matthew's Individual

Matt's Question: How does a longer gestation period affect the baby?

```{r}
LM <- lm(bwtoz ~ gestation, data = babies)
summary(LM)
coef(LM)

ggplot(data = babies,aes(y = bwtoz, x = gestation))+
  geom_jitter(na.rm = TRUE)+
  geom_smooth(method = "lm", na.rm = TRUE, color = "red")+
  xlab("Gestation in Days")+
  ylab("Baby Weight in Oz")+
  ggtitle("Babies Weights V.S. The Gestation period\nand the Linear Model that Shows the Trend")
```

Matt's Findings: I found with plotting the gestation time and babies birthweight on a scatter plot that there wasn't much to determine from it. But with the red line representing the linear model, it is easy to see the relationship between baby weight and gestation period. The line indicates that with more time in the womb of the mother the bigger the baby will be.  
Looking at the summary of the linear model, there is a very high degree of freedom being 1026 degrees which is exceptionally high. Due to the large sample size we were given with this data set the high degree of freedom means that we can reject the null hypothesis of the data and are more likely to find a strong mean for the data points. 


#### Shreeya's PLot

Shreeya's Question: Does the mother's age affect gestation period?

```{r}
babies_young_mothers<- babies %>% filter(mage <= 25)
babies_old_mothers<- babies %>% filter(mage >= 30)
#View(babies2)
babysim_mod <- lm(gestation ~ mage, data = babies)
grid <- babies %>% 
  data_grid(mage) 
grid <- grid %>% 
  add_predictions(babysim_mod) 
ggplot(data = babies) +
    geom_point(mapping=aes(x=mage, y =gestation)) + 
    geom_line(aes(y = pred, x = mage), data = grid, colour = "red", size = 1)+
    ggtitle("Does the Mother's Age affect the Gestation Period?") +
    xlab("Mother's Age") +
    ylab("Gestation Period")
```

Shreeya's Findings: This question is important to possibly eliminate a confounding factor to the question "Does smoking affect babies gestation periods?" My findings were that mother's age doesn't impact gestation period, thus eliminating a confounding variable. Looking at the linear model fit to this graph, it reinforces there is no trend between these variables. I used simple modeling tools to create all the parts of this plot.


### Team Summary

* I, Kelsey McKenna, created a plot that showed the relationship between the number of cigarettes smoked per day and the child's birthweight. I thought this was an important question and I had predicted that the more a mother smoked, the worse her health and the lower the birthweight of the child. It is clearly not the case and there is little relationship between the two. I used the geom_point function. I added a color function with the variable smoke to highlight if the mother said she smoked or not. I found that there are more higher weighted babies as outliers if the mother doesn't smoke. I added labels to make the plot easier to read. I also created the team plot number 1 and the question and analysis for our first team plot. 

I was able to modify my work, by making a better plot for the team one and filtering out certain values for whether a mother smoked at all. Doing showed showed that our results were inconclusive and not statistically significant. Therefore, we can not decide that a women smoking impacts their babies birthweight. More tests would need to be done.  

* I, Shreeya Basrur, created a scatterplot to analyse if there is a relationship between gestation period and mother's age. To answer this question, I created a simple linear model of the data I was looking at by first using the formula of a linear model, then creating a data grid of the mother's age data and adding predictions of y values. Finally, I plotted that on the graph of age versus gestation period to confirm a lack of a trend. 

* I, James Burke, wrote the first question (smoking vs premature babies) and its importance in warning mothers of the affect smoking can have on their newborns. I also helped to create and format the Number of Births vs. Number of Cigarettes Smoked Per Day plot. For my individual plot I created a bar chart that compares the proportion of smokers vs non smokers, to they income grouping. I made incomne the x axis and then had separate bars in each income grouping, for peoples smoking status. Having these bars side by side makes it easy to visualize the proportion. I did this using the fill aestetic and by making the smokers variable a factor, then I used position dodge to separate the bars. I also changed the x axis label along with the legend title.

* I, Matthew Copeland, created a graph looking at the gestation period compared to the baby birth weight. I first added the scatter plot and then the linear relationship between the two. In my mind i figured it could be more of a clear relationship with the scatter plot but the data showed other wise. I also added the summary of the linear model give some extra data to look at. 





























